### 2025-06-20 modern-portfolio-theory.qmd
20250620 - sum of players' variances

plot(
  efficientFrontierData$volatility,
  efficientFrontierData$return,
  pch = 19,
  cex = 0.5,
  type = "o",
  xlab = "Variance in Projected Fantasy Points Across Sources",
  ylab = "Projected Fantasy Points")
```

@@

@@

plot(
  efficientFrontierData2$volatility,
  efficientFrontierData2$return,
  pch = 19,
  cex = 0.5,
  type = "o",
  xlab = "Variance in Fantasy Points Across Weeks",
  ylab = "Weekly Fantasy Points")
```

>>>>

plot(
  efficientFrontierData$volatility,
  efficientFrontierData$return,
  pch = 19,
  cex = 0.5,
  type = "o",
  xlab = "Standard Deviation of Projected Fantasy Points Across Sources",
  ylab = "Projected Fantasy Points")
```

The formula for the variance (and standard deviation) of the sum of two variables is in @eq-varianceSumTwoVars:

$$
\begin{aligned}
  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)\\
  s^2_{(x + y)} &= s^2_x + s^2_y +  2\text{Cov}(x, y)\\
  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y) + 2\text{Cov}(x, y)}\\
  s_{(x + y)} &= \sqrt{s^2_x + s^2_y + 2\text{Cov}(x, y)}\\
\end{aligned}
$$ {#eq-varianceSumTwoVars}

If we assume that players' performance is uncorrelated with one another, the covariance between the players is zero, so the formula simplifies to @eq-varianceSumTwoVarsNoCov:

$$
\begin{aligned}
  \text{Var}(x + y) &= \text{Var}(x) + \text{Var}(y)\\
  s^2_{(x + y)} &= s^2_x + s^2_y\\
  \text{Std Dev}(x + y) &= \sqrt{\text{Var}(x) + \text{Var}(y)}\\
  s_{(x + y)} &= \sqrt{s^2_x + s^2_y}\\
\end{aligned}
$$ {#eq-varianceSumTwoVarsNoCov}

That is, if players' performance is independent of each other, the variance of two players' points is merely the sum of their variances; their standard deviation is then the square root of that.
In reality, players' performance is not truly independent—particularly for players on the same team or, for a given game, for players on opposing teams.
However, for players not playing in the same game, it is a reasonable assumption to make and it greatly simplifies the math.
If you wanted to, you could account for players' covariance by using @eq-varianceSumTwoVars.
An example variance-covariance matrix of players' performance is in @sec-efficientFrontierFantasyProjectionsGameToGameVariability.

@@

Here is the variance-covariance matrix of players' performance (it is very large because there are many players):

```{r}
dataForVarCovMatrix <- player_stats_weekly_recent %>% 
  select(player_id, week, fantasyPoints) %>% 
  pivot_wider(
    names_from = player_id,
    values_from = fantasyPoints
  )

cov(
  dataForVarCovMatrix,
  use = "pairwise.complete.obs")
```

@@

plot(
  efficientFrontierData2$volatility,
  efficientFrontierData2$return,
  pch = 19,
  cex = 0.5,
  type = "o",
  xlab = "Standard Deviation of Fantasy Points Across Weeks",
  ylab = "Weekly Fantasy Points")
```
